---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# (temporary name) dynIRTtest

<!-- badges: start -->
<!-- badges: end -->

We will change the package name later. 
Please edit the `README.Rmd` file and do not directly edit the `README.md` file.

## Installation

You can install the development version of **dynIRTtest** from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tomoya-sasaki/dynIRTtest")
```

## TODOs

- [x] Add descriptions
- [ ] Figure specifications

## Explanation

The R package **dynIRTtest** fits dynamic multidimensional factor models with
binary, ordinal, and/or metric indicators. To do so, it uses the Bayesian
programming language Stan, as linked to R by the package **CmdStanR**.

The basic workflow involves the following steps:
  1. Shape the data into the list format required by **CmdStanR**.
  2. Fit a dynamic factor model to the data.
  3. Extract draws from the fitted model.
  4. 

### Step 1: load data

Load data on state societal outcomes from 2020 and 2021.

```{r, eval=FALSE}
data("social_outcomes_2020_2021")
```

### Step 2: Reshape data

```{r, eval=FALSE}
shaped_data <- shape_data(
    long_data = social_outcomes_2020_2021,
    unit_var = "st",
    time_var = "year",
    item_var = "outcome",
    value_var = "value",
    standardize = TRUE,
    periods_to_estimate = 2020:2021
)

```

### Step 3: Fit the model

```{r,eval=FALSE}
fitted <- fit(
    data = shaped_data,
    lambda_zeros = data.frame(
        item = c("x_cps_spm_poverty"),
        dim = c(2)
    ),
    n_dim = 2,
    chains = 2,
    parallelize_within_chains = TRUE,
    threads_per_chain = 2,
    constant_alpha = FALSE,
    separate_eta = TRUE,
    init_kappa = FALSE,
    force_recompile = FALSE,
    iter_warmup = 50,
    iter_sampling = 50,
    adapt_delta = .9,
    refresh = 10,
    seed = 123
)
```

### Step 4: Post analysis

```{r,eval=FALSE}

```

